\textbf{ Spectrophotometric parallaxes with linear models:\\ Accurate distances for luminous red-giant stars }

### 2018-09-30 paper/sp.tex
upadte paper

% To-do items
% -----------
% - Coryn's point that we should carefully explain how our model is geometric.
% - check with WISE experts that we have the correct WISE photometry
% - write words about other linear models we have made or could make, probably in the discusssion.
% - are we citing all the proper DR2 papers we need to cite?
% - HWR: Footnote to explain the use of the word ``feature'' for astronomers.
% - HWR: Figures 1 and 2 need to be combined to directly compare Parent and Training.
% - HWR: Figure 3: Note that the y-axis is positive-definite!
% - HWR: Figure 7: Caption needs to say what arrows represent.

@@

\begin{abstract}\noindent
% context
With contemporary infrared spectroscopic surveys like \apogee,
red-giant stars can be observed to distances and extinctions
at which \gaia\ parallaxes are not informative.
Yet the combination of effective temperature, surface gravity, composition, and age---all
accessible through spectroscopy---determines a giant's luminosity.
Therefore spectroscopy plus photometry should enable 
precise spectrophotometric distance estimates.
% aims
Here we use the \apogee--\gaia\ overlap to train a data-driven model
to predict parallaxes for red-giant branch stars with $0<\logg<2.2$.
% methods
We employ (the exponentiation of)
a linear function of \apogee\ spectral pixel intensities and multi-band photometry
to predict a parallax spectrophotometrically.
The data-driven-model training involves no logarithms or inverses of the \gaia\ parallaxes,
and needs no cut on the \gaia\ parallax signal-to-noise ratio.
It includes an L1 regularization to zero out the contributions of
uninformative pixels.
The training is performed
with leave-out subsamples such that no star's astrometry is used even indirectly in its
spectrophotometric parallax estimate.
% results
The model implicitly performs a reddening and extinction correction in its parallax prediction,
without any explicit dust model.
We assign to each star in the sample a new spectrophotometric parallax estimate;
these parallaxes have uncertainties of a few to 15 percent, depending on data quality,
which is more precise than the \gaia\ parallax for the vast majority of targets.
We can now get 10-percent distance estimates out to heliocentric distances of 20\,kpc,
allowing truly global maps of the Milky Way's disk.
\end{abstract}

@@

As we operate these spectroscopic surveys,
the \gaia\ Mission (\citealt{gaia}) has been revolutionizing our view of the disk.
It delivers good proper motions everywhere,
and extremely valuable parallax information
locally.
However, it does not deliver precise geometric parallaxes over a large fraction of
the disk, and certainly not in the dusty and crowded regions.
For these reasons, the value of \gaia\ in these infrared spectroscopy projects is not
to deliver distance information directly, but rather to calibrate stellar models,
which then deliver distance information through relationships between stellar
luminosities and spectral characteristics.

@@

>>>>

% To-do items
% -----------
% - Coryn's point that we should carefully explain how our model is geometric.
% - check with WISE experts that we have the correct WISE photometry
% - write words about other linear models we have made or could make, probably in the discusssion.
% - are we citing all the proper DR2 papers we need to cite?
% - HWR: Footnote to explain the use of the word ``feature'' for astronomers.
% - HWR: Figures 1 and 2 need to be combined to directly compare Parent and Training.
% - HWR: Figure 3: Note that the y-axis is positive-definite!
% - HWR: Figure 7: Caption needs to say what arrows represent.
% - Make the comparison with Cannon as a function of L1 strength and maybe look at correlation coefficients between parameter vectors.
% - Get model running on ASPCAP outputs and compare in Discussion.
% - Make sure we incorporate / respond to Ness comments from Aug 13 email.
% - Thank APOGEE telecon and add paragraph about AGB stars and the color-magnitude diagram in the discussion.
% - Describe how we ran The Cannon; discussion of how we compare to The Cannon.
% - Cite and describe Eilers et al.

@@

\begin{abstract}\noindent
% context
With contemporary infrared spectroscopic surveys like \apogee,
red-giant stars can be observed to distances and extinctions
at which \gaia\ parallaxes are not informative.
Yet the combination of effective temperature, surface gravity, composition, and age---all
accessible through spectroscopy---determines a giant's luminosity.
Therefore spectroscopy plus photometry should enable 
precise spectrophotometric distance estimates.
% aims
Here we use the \apogee--\gaia\ overlap to train a data-driven model
to predict parallaxes for red-giant branch stars with $0<\logg\leq2.2$.
% methods
We employ (the exponentiation of)
a linear function of \apogee\ spectral pixel intensities and multi-band photometry
to predict a parallax spectrophotometrically.
The model training involves no logarithms or inverses of the \gaia\ parallaxes,
and needs no cut on the \gaia\ parallax signal-to-noise ratio.
It includes an L1 regularization to zero out the contributions of
uninformative pixels.
The training is performed
with leave-out subsamples such that no star's astrometry is used even indirectly in its
spectrophotometric parallax estimate.
% results
The model implicitly performs a reddening and extinction correction in its parallax prediction,
without any explicit dust model.
We assign to each star in the sample a new spectrophotometric parallax estimate;
these parallaxes have uncertainties of a few to 15 percent, depending on data quality,
which is more precise than the \gaia\ parallax for the vast majority of targets.
We can now get 10-percent distance estimates out to heliocentric distances of 20\,kpc,
allowing truly global maps of the Milky Way's disk.
\end{abstract}

@@

As we operate these spectroscopic surveys,
the \gaia\ Mission (\citealt{gaia}) has also been revolutionizing our view of the disk.
It delivers good proper motions everywhere,
and extremely valuable parallax information locally.
However, it does not deliver precise geometric parallaxes over a large fraction of
the disk, and certainly not in the dusty and crowded regions.
For these reasons, the value of \gaia\ in these infrared spectroscopy projects is not
to deliver distance information directly, but rather to calibrate stellar models,
which then deliver distance information through relationships between stellar
luminosities and spectral characteristics.

@@

ACE: ADD HERE the Color-magnitude diagrams for high snr and training set.
HOGG: DIscussion here of CMD and AGB stars.
